Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-07-21 Document/paper.tex
test

% Abstract of the paper
\begin{abstract}
We present shape measurements of Milky Way sized dark matter halos at
redshift $z=0$ in a suite of 30 zoom simulations from the Auriga
project. 
We compare the results in full magnetohydrodynamics against dark
matter only physics. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
At distances $\lesssim 30$ kpc, rounder dark matter distributions
correlate with extended massive stellar discs and low gas 
densities.  
We measure the alignment between the halo and the stellar disc and
find almost perfect alignment at $0.25R_{200}\sim56$ kpc.   
In some cases the alignment between the two components significantly
changes as a function of radius implying that the halo shape twists;
this effect correlates with extended stellar discs and
low gas densities and is almost absent in the dark matter only
simulations.  
In a comparison against observational constraints  we find that $20\%$
of halos are consistent with observational results derived
from the Pal 5 stream that favor an almost spherical shape.
Including baryons is a required element to achieve this
level of agreement. In contrast, none of the simulations (dark matter
only nor with baryons) match the constrains derived from the
Sagittarius stream favoring an oblate dark matter halo.
\end{abstract}

@@

We start the calculations taking into account particles within a
sphere of radius $R$ and then recharacterize the triaxial parameters
by taking into account particles within an ellipsoid of semi-axes
$r,r/q,r/s$ and re-scaled distance $d^2=x^2+(y/q)^2+(z/s)^2$, where $q
= b/a$ and $s=c/a$ are the previously calculated axial ratios. 
We repeat this process until the average deviation of semi-axes is
less than $10^{-6}$.  
After convergence we define a unique radius $R$ as the geometrical
mean of the axial lenghts $R=(abc)^{1/3}$.
We use this radial coordinate $R$ to parameterize the spatial changes
in halo shape we report in the following sections.

This is the same method used to estimate the halo shape in the DM-only
Aquarius simulations \citep{VeraCiro11}.

Following the convergence criterion by \cite{VeraCiro11} we
restrict the sampling of the ellipsoidal parameters to radii  between
$\sim 2$kpc and $R_{200}$, where  $R_{200}$ correspond to the  radius
enclosing a sphere with 200 times the critical density of the Universe.
On average, over the 30 halos in the \texttt{Level4} sample
$R_{200}=230\pm 15$kpc.

@@

In this paper we measured the shape of 30 isolated Milky Way like 
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
The Auriga project simulated these halos with two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
including star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these two kinds of simulations to quantify the
halo shape as a function of radius and the degree of aligment between
the stellar disc and the halo.

We find that MHD halos are rounder than DMO halos at every radius.
MHD halos tend towards more oblate shapes ($T<1/3$), while DMO halos
tend towards more prolate shapes ($T>2/3$).  
The rounding effect is more noticeable as one moves closer to galactic
disc and strongly  correlates with baryonic properties of the disc
More precisely, the triaxility is smaller for large and massive
stellar discs with low gas densities at its core.

We also measured the shape alignment of the halo with the stellar
disc.
At a radius of $0.25R_{200}$ the alignment is almost perfect and
degrades at smaller and larger distances.
We found that in some halos the alignment betwen the stellar disc
and dark matter halo changes noticeably with the radius.
This alignment evolution implies a radial twisting between the ellipsoids
describing the halo shape. 
We quantify this twist with the standard deviation of the angle
between the two minor axis at radii below $\leq 0.5R_{200}$ and find
that it strongly correlates with the disc size and the core gas
mass density, there is only a weak correlation with the disc to
total mass ratio in the disc.

We compared our results against two observational constraints for the
dark matter halo shape of the Milky Way. 
The constraints are at two different radii and come from different
observatinal tracers. 
We find that $20\%$ halos in the MHD simulations are consistent with
the constraints by \cite{Bovy16} at $\approx 19$kpc, while none of the
halos, either in MHD or DMO, has some overlap with the shape
constraints by \cite{LM10} at $\approx 40$kpc.

@@

From the purely numerical point of view a better understading of the
disc-halo alignment details will certainly require to study how the
halo and the disc co-evolved as a function of time.
Calibrating the effect of the cosmic web (both dark matter and gaseous)
\citep{2014MNRAS.443.1090F,2017MNRAS.469..594B,2019MNRAS.487.1607G}
could also be required to gain a full picture of this behaviour.

>>>>

% Abstract of the paper
\begin{abstract}
We present shape measurements of Milky Way sized dark matter halos at
redshift $z=0$ in a suite of 30 zoom simulations from the Auriga
project. 
We compare the results in full magnetohydrodynamics against dark
matter only simulations. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
At distances $\lesssim 30$ kpc, rounder dark matter distributions
correlate with extended massive stellar discs and low core gas 
densities.  
We measure the alignment between the halo and the stellar disc at
different radii and find almost perfect alignment at
$0.25R_{200}\sim56$ kpc.    
In some cases the alignment significantly changes as a function of
radius implying that the halo shape twists; 
this effect correlates with extended stellar discs and
low gas densities and is almost absent in the dark matter only
simulations.  
In a comparison against observational constraints we find that $20\%$
of halos in our sample are consistent with observational results derived
from the Pal 5 stream that favor an almost spherical shape.
Including baryons is a required element to achieve this
level of agreement. In contrast, none of the simulations (dark matter
only nor with baryons) match the constrains derived from the
Sagittarius stream favoring an oblate dark matter halo.
\end{abstract}

@@

We start the calculations taking into account particles within a
sphere of radius $R$ and then recharacterize the triaxial parameters
by taking into account particles within an ellipsoid of semi-axes
$r,r/q,r/s$ and re-scaled distance $d^2=x^2+(y/q)^2+(z/s)^2$, where $q
= b/a$ and $s=c/a$ are the previously calculated axial ratios. 
We repeat this process until the average deviation of semi-axes is
less than $10^{-6}$.  
After convergence we define a unique radius $R$ as the geometrical
mean of the axial lenghts $R=(abc)^{1/3}$.
We use this radial coordinate $R$ to parameterize the spatial changes
in halo shape we report in the following sections.
This is the same method used to estimate the halo shape in the DM-only
Aquarius simulations \citep{VeraCiro11}.

Following the convergence criterion by \cite{VeraCiro11} we
restrict the sampling of the ellipsoidal parameters to radii between
$\sim 2$kpc and $R_{200}$, where  $R_{200}$ correspond to the radius
enclosing a sphere with 200 times the critical density of the Universe.
On average, over the 30 halos in the sample $R_{200}=230\pm 15$kpc. 
In our results we express the radius as a fraction of $R_{200}$ in
order to have a consistent radius scale across all halos.

@@

In this paper we measured the shape of 30 isolated Milky Way sized
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
The halos were with two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
including star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these simulations to quantify the halo shape as a
function of radius and the degree of aligment between the stellar disc
and the halo.

We find that MHD halos are rounder than DMO halos at every radius.
MHD halos tend towards more oblate, sometimes almost spherical, shapes
($T<1/3$), while DMO halos tend towards more prolate shapes ($T>2/3$).  
The rounding effect is more noticeable as one moves closer to galactic
disc and strongly  correlates with baryonic properties of the disc
More precisely, the triaxility is smaller for extended and massive
stellar discs with low gas densities at its core.

We also measured the shape alignment of the halo with the stellar
disc at different radii.
At a radius of $0.25R_{200}$ the alignment is almost perfect and
degrades at smaller and larger distances.
In some halos the alignment betwen the stellar disc and dark matter
halo changes noticeably with radius. 
This alignment evolution implies a radial twisting between the ellipsoids
describing the halo shape. 
We quantify this twist with the standard deviation of the angle
between the two minor axis at radii below $\leq 0.5R_{200}$ and find
that the twisting strongly correlates with the disc size and the core
gas mass density, there is only a weak correlation with the disc to
total mass ratio.

We compared our results against two observational constraints for the
dark matter halo shape of the Milky Way. 
The constraints are at two different radii and come from different
observatinal tracers. 
We find that $20\%$ halos in the MHD simulations are consistent with
the constraints by \cite{Bovy16} at $\approx 19$kpc that correspond to
an almost spherical halo, while none of the halos, either in MHD or
DMO, has some overlap with the shape constraints by \cite{LM10} at
$\approx 40$kpc that is closer to oblate.

From the purely numerical point of view, a better understading of the
influence of baryons on the different properties we have measure  will
certainly require to study how the halo and the disc co-evolved as a
function of time. 
\cite{VeraCiro11} advanced suggested the idea that the current dark
matter halo shape strongly correlates with the time evolution of the
halo as traced by the shape measured at the virial radius.
The opposed trend of triaxiality as a functio of radius between the
MHD and the DMO simulations, together with twsiting effect in some of
the halos seems to suggest that this effect might no longer hold.

Calibrating the effect of the cosmic web (both dark matter and gaseous)
\citep{2014MNRAS.443.1090F,2017MNRAS.469..594B,2019MNRAS.487.1607G}
could also be required to gain a full picture of this behaviour.

@@
